Post-Disturbance Tree Community Trajectories in a Neotropical Forest

### 2018-02-19 WholePlotTrajectories.Rmd
Manuscript update

Our study underlined the significant impact of disturbance on tropical forests communities but a contrasting response of taxonomic and functional diversity.
Specifically,  it unveiled the decoupling between trajectories of taxonomic evenness and their functional diversity and dominant characteristics.
This decoupling confirms the independence for tropical forests between the processes grasped through the studied functional traits (primary productivity, water and nutrient cycling, etc.) and communities composition and evenness as it was suggested .  
This highlighted the functional overlap between species in tropical forests,typical of their huge biodiversity, and intricately questioned the different facets of communities resilience.

@@

Although recovery seemed on its way the trajectories revealed significant changes in the ecological rules shaping communities over the 30 years of survey.
Disturbance is thought to shift exclusive competition favoring competitive species to environmental filtering selecting acquisitive and disturbance resistant species [@TerSteege2001;@Flores2006;@Haddad2008]
The comparison of observed trajectories to a null model with random traits set shuffling showed an oriented functional shift after disturbance which suggest additional biotic and abiotic filters for the 10 to 15 years after disturbance.
In particular this is demonstrated by the evolution of seed mass classes proportions revealing a clear increase in small seeded and then large dispersive species.
Faster than for taxonomic composition and diversity though, functional characteristic returned to their initial values and the values observed for control plots.
This suggested some functional overlap in the pre-disturbance condition (*i.e* the usual tropical forests conditions). 
Such overlap and maybe functional versatility was already observed in highly diverse ecosystems [@Bellwood2006] and sugegsted as one mechanisms for high diversity maintenance.
Decoupled taxonomic and functional trajectories questions to what extent communities recovery is completed: ecosystem functioning might be quickly recoverd, and taxonomic composition following suit although along for much longer, but the resilience of the community itself might be altered as functional overlap themselves would be lost.
Along with probably other environmental modifications (like soils nutrient cycling and compaction [@Olander2005]) with alteration of communities resilience makes us expect different trajectories after subsequent disturbance.
As disturbance seemed to update communities toward more disturbance resistant ones [@Haddad2008] it is possible that eventually the disturbance regie observed has set another permanent condition, which would only be more diverse if good competitors, poor disturbance resitant survive.
> (Flores, 2006  for differential sapling success after disturbance)

>>>>

Our study underlined the significant impact of disturbance on tropical forests communities but a contrasting response of taxonomic and functional diversity.
Specifically,  it unveiled the decoupling between trajectories of taxonomic evenness and their functional diversity and dominant characteristics.
This decoupling confirms the "vegetation quantity effect" suggesting that communities functional characteristics are determined by dominant species rather that subtle changes in biodiversity [@reftobefound]. This would result in a faster recovery of ecosystem processes rates than in communities ocmposition and diversity as it mas largely demonstrated for grasslands [@Tilman1997; @Mokany2008; @Mouillot2011 *in Lohbeck2015*] and recently proved for tropical forests [@Lohbeck2015].  
This highlighted the functional overlap between species in tropical forests,typical of their huge biodiversity, and intricately questioned the different facets of communities resilience.
Ecosystem processes graped through the set of functional traits retained here however do not assess communities resilience and future dynamics.

@@

Although recovery seemed on its way the trajectories revealed significant changes in the ecological rules shaping communities over the 30 years of survey.
Disturbance is thought to shift exclusive competition favoring competitive species to environmental filtering selecting acquisitive and disturbance resistant species [@TerSteege2001;@Flores2006;@Haddad2008]
The comparison of observed trajectories to a null model with random traits set shuffling showed an oriented functional shift after disturbance which suggest additional biotic and abiotic filters for the 10 to 15 years after disturbance.
In particular this is demonstrated by the evolution of seed mass classes proportions revealing a clear increase in small seeded and then large dispersive species.
Faster than for taxonomic composition and diversity though, functional characteristic returned to their initial values and the values observed for control plots.
This suggested some functional overlap in the pre-disturbance condition (*i.e* the usual tropical forests conditions). 
Such overlap and maybe functional versatility was already observed in highly diverse ecosystems [@Bellwood2006] and sugegsted as one mechanisms for high diversity maintenance.
Decoupled taxonomic and functional trajectories questions to what extent communities recovery is completed: ecosystem functioning might be quickly recoverd, and taxonomic composition following suit although along for much longer, but the resilience of the community itself might be altered as functional overlap themselves would be lost.
Along with probably other environmental modifications (like soils nutrient cycling and compaction [@Olander2005]) with alteration of communities resilience makes us expect different trajectories after subsequent disturbance.
As disturbance seemed to update communities toward more disturbance resistant ones [@Haddad2008] it is possible that eventually the disturbance regime observed has set another permanent condition, which would only be more diverse if good competitors, poor disturbance resitant survive.
We assume that such post-disturbance forest community might have the same functioning, as suggested by the functional trajectories, but composition displays a poor recovery, especially regarding species contingent to undisturbed forests or different life forms as lianas or epiphytes that have not been monitored here [@Martin2013].
Although forests functioning and processes seemed likely to recover after the range of disturbance studied, the hampered recovery of communities composition might jeopardize the conservation of some species, therefore threatening the global biodiversity [@Gardner2007].
Additionally to threats on particular species conservation the consequences of biodiversity changes on hyper-diverse ecosystems might reveal dramatic, specifically in impairing ecosystems resilience and resistance to further disturbance [@Trenbath1999;@Elmqvist2003;@Diaz2006].
